%%% ============================================================
%%% Quantum-Enhanced Weather Classification: SVMs with IQP
%%% Encoding for Cyclone Prediction
%%% ============================================================
%%% 
%%% Target Journal: Springer EPJ Quantum Technology or
%%%                 Quantum Information Processing
%%%
%%% INSTRUCTIONS:
%%% 1. Download Springer Nature's LaTeX template from:
%%%    https://www.springernature.com/gp/authors/campaigns/latex-author-support
%%% 2. Place sn-jnl.cls and sn-mathphys-num.bst in this directory
%%% 3. Compile with: pdflatex -> bibtex -> pdflatex -> pdflatex
%%%
%%% If you do not have the Springer class file, uncomment the
%%% fallback line below and comment out the sn-jnl line.
%%% ============================================================

%% --- Springer Nature Journal Class ---
%% Uncomment ONE of the following:
\documentclass[sn-mathphys-num,Numbered]{sn-jnl}
%% \documentclass[12pt,a4paper]{article}  % Fallback if sn-jnl.cls unavailable

%%% ============================================================
%%% PACKAGES
%%% ============================================================
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{graphicx}
\usepackage{geometry}
\geometry{hmarginratio=1:1}  % Centre text block on all pages
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{tikz}
\usetikzlibrary{quantikz2}  % For quantum circuit diagrams
\usetikzlibrary{arrows.meta,positioning,shapes.geometric,calc,fit,backgrounds}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usepackage{xcolor}
% \usepackage{microtype} % Disabled to prevent 'auto expansion is only possible with scalable fonts' error

%%% ============================================================
%%% METADATA
%%% ============================================================
%% \jyear{2026}  % Uncomment if your sn-jnl.cls version supports this

\title[Quantum-Enhanced Weather Classification]{%
Quantum-Enhanced Weather Classification: Support Vector Machines 
with IQP Encoding for Cyclone Prediction}

%% --- Authors ---
\author*[1]{\fnm{Parthiban} \sur{K}}\email{k.parthiban@vit.ac.in}
\author[1]{\fnm{Bhavya} \sur{Dhoot}}\email{bhavya.dhoot@vit.ac.in}

\affil*[1]{\orgdiv{School of Computer Science and Engineering}, 
\orgname{Vellore Institute of Technology}, 
\orgaddress{\city{Vellore}, \postcode{632014}, \state{Tamil Nadu}, \country{India}}}

%%% ============================================================
%%% CUSTOM COMMANDS
%%% ============================================================
\renewcommand{\ket}[1]{|#1\rangle}
\renewcommand{\bra}[1]{\langle#1|}
\renewcommand{\braket}[2]{\langle#1|#2\rangle}
\newcommand{\expect}[1]{\langle#1\rangle}
\newcommand{\norm}[1]{\left\|#1\right\|}
\newcommand{\abs}[1]{\left|#1\right|}
\newcommand{\tr}{\operatorname{Tr}}
\newcommand{\iqp}{\textsc{iqp}}
\newcommand{\qsvm}{\textsc{qsvm}}

%%% ============================================================
%%% DOCUMENT
%%% ============================================================
\begin{document}

\input{sections/abstract}

\maketitle

\input{sections/introduction}
\input{sections/related-work}
\input{sections/background}
\input{sections/methodology}
\input{sections/experiments}
\input{sections/discussion}
\input{sections/conclusion}

%%% ============================================================
%%% ACKNOWLEDGEMENTS
%%% ============================================================
\begin{acknowledgements}
The authors acknowledge the use of IBM Quantum services and the Qiskit software framework for quantum circuit simulation. We also thank the National Oceanic and Atmospheric Administration (NOAA) for making the International Best Track Archive for Climate Stewardship (IBTrACS) dataset publicly available. Computational resources were provided by Vellore Institute of Technology.
\end{acknowledgements}

%%% ============================================================
%%% DECLARATIONS (EPJ Quantum Technology format)
%%% ============================================================

\noindent\textbf{Author contributions}
\par\noindent
P.K. conceived the study, supervised the research, and reviewed the manuscript. B.D. designed and implemented the quantum kernel framework, conducted the experiments, analysed the results, and wrote the manuscript. Both authors read and approved the final version.

\bigskip
\noindent\textbf{Funding}
\par\noindent
Open access funding provided by Vellore Institute of Technology. Not Applicable.

\bigskip
\noindent\textbf{Data Availability}
\par\noindent
All datasets used in this study are publicly available. The IBTrACS dataset is accessible from NOAA at \url{https://www.ncei.noaa.gov/products/international-best-track-archive}.

\bigskip
\noindent\textbf{Code availability}
\par\noindent
Source code for the quantum kernel SVM implementation is available at \url{https://github.com/Bhavya-Dhoot/quantum-cyclone-classification.git}.

\bigskip
\noindent\textbf{Declarations}

\noindent\textbf{Competing interests}
\par\noindent
The authors declare no competing interests.

%%% ============================================================
%%% BIBLIOGRAPHY
%%% ============================================================
\bibliographystyle{sn-mathphys-num}
\bibliography{references}

\end{document}
