%%% ============================================================
%%% SECTION 6: DISCUSSION
%%% ============================================================
\section{Discussion}\label{sec:discussion}

The preceding mathematical performance metrics establish an empirical baseline for IQP-based quantum kernels processing complex atmospheric data arrays. Accuracy tables require isolated contextual analysis. This analysis parses specific model constraints, gross computational overhead, and explicit NISQ hardware limitations directly.

\subsection{Interpretation of Results}

The raw output data reveals a stark classification capability split. The QSVM-IQP evaluating depth $L = 1$ secured 96.1\% global accuracy. This model processed a heavily restricted 600-sample operational training slice. Classical software baselines trained on ten times the raw data volume (5{,}493 specific samples). These classical models reached approximate 100\% validation accuracy. This massive dataset asymmetry drives the explicit performance gap. Classical SVM architecture successfully identified fundamental linear separability rules within the six-variable atmospheric feature space. This capability depended entirely upon the provision of overwhelming coordinate training volume.

Quantum model analysis requires precise comparisons evaluating algorithms under identically constrained data parameters. The IQP feature map structured at $L = 1$ hit an isolated 96.1\% diagnostic accuracy ceiling. The comprehensively entangled ZZFeatureMap baseline recorded exactly 83.9\% accuracy. Generic entanglement algorithms provided absolutely no structural classification benefit for this specific meteorological mathematics problem. The built IQP map pairs native explicit single-variable encoding directly against targeted pairwise mathematical products. This distinct topological structural protocol proved vastly superior.

The per-class performance data table (Table~\ref{tab:per-class}) isolates the exact physical source of this algorithmic advantage. The IQP mathematical model defeated the generic ZZ map architecture most significantly within the moderate hurricane bracket (Class~1). This specific intensity class represents the most complex mathematical classification boundary globally. A simulated storm sustaining exactly 64~knots borders both peak-level tropical storm and lowest-tier Category~1 hurricane metric classification vectors. Individual isolated wind and atmospheric pressure readings project identically across this narrow data boundary. Latitude positioning, chronological translational speed, and physical radius of maximum internal winds act necessarily as primary disambiguating feature variables. The IQP hardware circuit utilises controlled-phase gates natively. These gates imprint specific physical joint-effect interactions directly into the calculated kernel map matrix. Generic dense entangling configurations failed completely to mirror this exact atmospheric physical coupling algorithm.

The kernel-target alignment calculations (Table~\ref{tab:alignment}) reinforce these core structural findings explicitly. The IQP output kernel scaling at $L = 1$ registered a mathematical alignment of 0.219. The comparative ZZFeatureMap managed a baseline 0.105 total alignment score. This disparity differential represents a rigorous mathematical efficiency improvement. Matrix alignment scores quantify exactly how accurately specific operational kernel geometries match targeted outcome label structures. The shallowest functional IQP kernel recorded the absolute highest entangled alignment capability metric. Extending explicit IQP processing depths directly to $L=2$ and $L=3$ degraded this geometric alignment calculation severely. Over-parameterised operational quantum feature spaces immediately generate explicit systemic geometric liabilities rather than yielding any structural classification validation advantages.

\subsection{Computational Considerations}

Simulating extensive quantum kernels mathematically via numerical statevector propagation models requires intensive hardware processing resources. Every individual single circuit evaluation loop involves multiplying isolated $2^n \times 2^n$ dense tracking matrices. The total assembly formulation calculating the full operational kernel matrix demands absolute $O(N^2)$ independent execution evaluations. Our designed six-qubit topological architecture maintains a specifically manageable $2^6 = 64$ dimensional Hilbert phase state space limit. Hardware parameters scaling upwards to twelve or fifteen distinct mapping qubits shift the primary total processing bottleneck directly over to the classical quantum simulator interface.

Generating the restricted diagnostic $600 \times 600$ test vector kernel matrices consumed incredibly heavy baseline CPU processor time limits. Simulation calculation processing scales explicitly quadratically mapped against total input dataset accumulation volume. The operational classical RBF kernel evaluated all distinct 5{,}493 isolated training parameter samples flawlessly in under one absolute physical processing second. Advancing core classical software optimisation frameworks cannot resolve this fundamental absolute computational capability gap architecture. Operational, real-world meteorological viability mandates functional physical quantum processing hardware access unconditionally. Functional Quantum Processing Units (QPUs) executing native structural IQP-type operating gates replace mathematical linear matrix multiplication entirely with direct physical electron measurement mechanics natively. Processing wall-clock operational execution runtimes will scale linearly against hardware circuit depth parameters and defined quantum shot counts rather than exploding outward into exponential Hilbert-space variable tracking dimensions.

\subsection{Limitations}

This algorithmic parameter investigation operates continually under four explicit mathematical experimental execution boundary constraints.

\paragraph{Noiseless simulation.} We gathered all core diagnostic calculation metric operational data exclusively via a perfected software statevector simulator loop. The mathematical evaluation background environment contained absolutely no induced measurement gate errors, registered zero physical qubit decoherence tracing limits, and experienced zero final output measurement statistical noise generation. Operating functional physical NISQ sector hardware displays vastly different uncorrected performance interaction parameters natively. Base two-qubit hardware gate fidelities fluctuate erratically between functional 99\% boundaries and ceiling 99.5\% measurement performance markers. Base qubit structural hardware relaxation algorithms ($T_1$) and uncorrected phase separation dephasing metrics ($T_2$) dictate maximum functional operational circuit longevity scaling spans inherently. Adjacent operating parameter gate algorithmic crosstalk degradation loops continuously erode and degrade precise kernel diagnostic geometric mapping topology estimates uncontrollably. Baseline operational error cancellation mitigation mathematical mechanisms exist presently~\cite{huang2021}. Systemic zero-noise calculation probability extrapolation and direct probabilistic geometric structural error cancellation routines represent standard current field interventions. Uncorrected base hardware operational transition mathematical survivability models for these highly specific precise performance classification accuracy structural metrics remain entirely empirically untested.

\paragraph{Restricted feature set.} Six constrained discrete meteorological recording parameters constitute a mathematically completely restricted system calculation input space boundary. Top-tier operational cyclone intensification diagnostic simulation models routinely deploy massively larger operational feature variable tracking sets. Calculating oceanic sea surface bounding temperatures, specific dynamic longwave radiation upper-atmosphere fluxes, absolute measurable vertical parameter wind shear gradient boundaries, and Convective Available Potential Energy (CAPE) scalar measurements all influence physical cyclone structural storm formation tracking dynamics inherently. Injecting these dense predictive variables dynamically into the protocol requires significantly expanding global qubit counts or executing complex destructive classical dimensionality reduction simplification steps initially. The IQP mathematical parameter kernel's discrete numerical tracking advantage algorithms may deteriorate immediately mapped natively against significantly expanding and multiplying complex global atmospheric scalar feature counts natively.

\paragraph{Agency biases in IBTrACS.} The master global IBTrACS operational database permanently merges localized individual best-track historical estimation records imported directly from independent dispersed international governmental meteorological administrative organisations. These individual isolated monitoring agencies structurally apply highly divergent internal mechanical intensity classification assessment estimation scaling protocols inherently. The compiled final global dataset permanently contains well-documented persistent systemic data input tracking biases inherently~\cite{landsea2013}. Regional output tracking data imported from Indian Ocean observation boundaries and combined Western Pacific isolated storm event records display continuous notable statistical tracking variance parameters natively. Restricting our analytical operational processing boundary dataset timeline to the specific post-2004 era spanning out until 2023 lessens, but absolutely does not entirely mathematically eradicate, these foundational historical observation scaling estimation discrepancies directly.

\paragraph{Fixed scaling parameters.} We manually hardcoded the exact IQP phase-map circuit definition operating scalar variables permanently isolating $\alpha_j = 1$ bound identically against $\beta_{jk} = 1$. This specific algorithmic circuit structural operational simplification restriction bypassed all functional quantum machine variable alignment optimization procedures. Employing iterative machine pretraining sequence feedback processing loops maximizing functional system kernel-target data alignment calculations could potentially mathematically discover superior isolated phase mapping tracking interaction angles securely. This specific optimization validation processing procedure natively demands significantly expanded computational supplementary software hyper-parameter algorithm processing overhead evaluation analysis budgets structurally.
